---
title: "simulate_cycle"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simulate_cycle}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
devtools::load_all()
library(VeloSim)
```
## Step 1: Define the simulation parameters
```{r params_setting}
# number of total cells
ncells_total <- 500
# number of total genes
ngenes <- 500
# total number of evfs: homo-evf + diff-evf
nevf <- 20
# number of diff-evf
n_de_evf <- 12
# which kinetic parameters is changed according to the sampling position, n_de_evf is used in s
vary <- "all"
# the standard deviation of the gaussian distribution, denote the noise size
Sigma <- 0.1
# the mean of the homo-evf, gaussian distribution again
evf_center <- 1
# with which probability the gene effect value is preserved
gene_effect_prob <- 0.3
# the mean of the gaussian distribution of gene effect
geffect_mean <- 0
# the standard deviation of the gaussian distribution of gene effect
gene_effects_sd <- 1
# the real data to obtain the kinetic parameter distribution from
param_realdata <- "zeisel.imputed"
# param_realdata <- NULL

bimod <- 0
scale_s <- 1
randseed <- 3
# number of cells in the second cycle
nextra <- 100
# number of cells to be removed at the beginning of the simulation
n_unstable <- 10
```

## Step 2: Simulate data according to the parameters
`SimulateVeloCycle` is the main function of VeloSim. This function is designed specifically for cell-cycle trajectory structure, user can generate the unspliced, spliced counts, true developmental time and true RNA velocity and kinetic parameters using this function.
```{r simulation}
# simulate data
result_true <- SimulateVeloCycle(ncells_total=ncells_total,ngenes=ngenes, evf_center=1,nevf=nevf, 
                            randseed=randseed, n_de_evf=n_de_evf,vary=vary,Sigma=Sigma,
                            geffect_mean=geffect_mean,gene_effects_sd=gene_effects_sd,
                            gene_effect_prob=gene_effect_prob,bimod=bimod,param_realdata=param_realdata,
                            scale_s=scale_s, prop_hge=0.015, mean_hge=5, nextra = nextra, n_unstable=n_unstable, plot = FALSE)
# add technical noise
result <- technicalNoise(result_true, capture.rate = 0.2)
```

## Step 3: Store the result
Store the unspliced count, spliced count, cell developmental time and RNA velocity
```{r save_file}
# unspliced count
write.table(result$counts_u, file = "./counts_u.csv")
# spliced count
write.table(result$counts_s, file = "./counts_s.csv")
# cell developmental time
write.table(result$cell_time, file = "./pseudo_time.csv")
# RNA velocity
write.table(result$velocity, file = "./velocity.csv")
```

## Step 4: Plot the result
```{r plotting}
# Plot with cell developmental time(pseudo-time), default PCA
plotPseudotime(filename = "pseudotime.pdf", result)
# Plot with RNA velocity
plotVelo(filename = "velocity.pdf", result, arrow.length = 0.3, width = 7, height = 7)
# Plot with cell developmental time(pseudo-time), using UMAP
plotPseudotime(filename = "pseudotime_umap.pdf", result, dr = "umap")
```
